Optimal strategies of influence in structured populations

### 2015-03-20 paper/manuscript.tex
small changes
added one line

We aim to assess optimal strategies assuming some cost associated with reducing the noise. Cost is per unit time and assumed to scale exponentially so that forcing $\sigma \rightarrow 0$ incurs exponentially increasing costs,
\begin{equation}
\mbox{Cost} = \tau (\sigma) e^{\beta N/\sigma}
\label{cost}
\end{equation}
Minimizing Eqn.~\ref{cost} means the optimal level of $\sigma$ is one for which
\begin{equation}
 \frac{d \ln \tau}{d \sigma} = N \beta \sigma^{-2}
\end{equation}
%\begin{figure*}
%\centering
%\includegraphics[scale=0.4]{figures/fig-chi.eps} 
%\caption{Probability to transition as a function of local neighbourhood for different values of the sociality weighting $\chi$}
%\label{fig1}
%\end{figure*}
We need to find $\tau$.
%\section{Mean first passage time away from incorrect state}
Consider the process as a one-step Markov chain on the integer line from $0$ to $N$. A transition from state $j$ to $j+1$ is written $t^+(j)$ and an opposite transition is denoted $t^-(j)$. An exact expression for the mean first passage time from zero individuals correct to $N_{ESC}$ is \cite{gardiner}

@@

%\section{Transition rates}
The dynamics of an update step in stochastic decision process are as follows; first an individual is selected at random from the population, this individual has a personal estimate of the environmental cue $g_i$ and observes the opinions of all its neighbors. The decision rule of Eqn.~\ref{opt_rule} is then employed and the individual will adopt, or remain in, the state most likely to be correct.
For any value of social information $x_L$ there is a value of personal information above which the individual will switch to $S=1$. We denote this value as $g_c$ and from Eqn.~\ref{opt_rule} 
\begin{equation}
g_c  = \sigma\ln\chi  \left(\frac{1}{2}-x_L\right) 
\end{equation}
If $g_i>g_c$ then the individual will switch to $S=1$ otherwise $S=-1$. 
Dealing first with the transition rate toward the correct state, we may write this as
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^+(x) = P(S_i=-1)\sum_{x_{L}}P\left(g_i>g_c \right)P\left(x_{L}|x\right)
\label{up}
\end{equation}
where $i$ denotes the selected individual and the summation is taken over all possible values of $x_{L}$.

From Eqn.~\ref{lapdist}
\begin{align}
P\left(g_i>g_c \right) =
\frac{1}{2}\exp{\left[1/\sigma - \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }.%&\mbox{if } \\
%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%\end{cases}
\end{align}
(This expression is true if $x_L \le 1/2 - 1/(\sigma\ln\chi)$. As we are interested in these probabilities only when $x$ is small we restrict our attention to this scenario.) Given this equation we may write Eqn.~\ref{up} as
\begin{equation}
t^+(x) = (1-x)  e^{1/\sigma } \mathcal{F}(x)
\end{equation}
where 
\begin{equation}
 \mathcal{F}(x) = \frac{1}{2}\sum_{x_{L}}\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x_L\right)\right]}P\left(x_{L}|x\right)
\end{equation}
and similarly 
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^-(x) =x \left(1-  e^{1/\sigma } \mathcal{F}(x)\right)
\end{equation}
Taking the derivative of Eqn.~\ref{laplace_cost} and employing a Maclaurin series expansion leaves
\begin{equation}
  \frac{d \ln \tau}{d \sigma} \approx - N \sigma^{-2} \int_{x_1}^{x_2} \left(1 + \sum_{n=1}^{\infty} \left(e^{1/\sigma } \mathcal{F}(y)\right)^n  \right) dy
\label{laplace_cost2}
\end{equation}
The extrema of the potential, $x_1,x_2$, are at the points where $t^+(x)=t^-(x)$, i.e. where $x=e^{1/\sigma } \mathcal{F}(x)$. Since $e^{1/\sigma } \mathcal{F}(x)<x$ $\forall x \in [x_1,x_2]$ the summation terms in the integrand are all strictly less than $x$, hence
\begin{equation}
  \frac{d \ln \tau}{d \sigma} \approx - N \sigma^{-2} \left({x_2}-{x_1}\right) + \mathcal{O}\left( \left(x_2-x_1\right)^2\right)
\label{laplace_cost2}
\end{equation}
%\begin{align}
%P\left(g_i>g_c \right) =
%1 - \frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }%&\mbox{if } \\
%%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%%\end{cases}
%\end{align}
%otherwise.

>>>>

We aim to assess optimal strategies assuming some cost associated with reducing the noise. Cost is per unit time and assumed to scale exponentially so that forcing $\sigma \rightarrow 0$ incurs exponentially increasing costs,
\begin{equation}
\mbox{Cost} = \tau (\sigma) e^{\beta N/\sigma}
\label{cost}
\end{equation}
Minimizing Eqn.~\ref{cost} means the optimal level of $\sigma$ is one for which
\begin{equation}
  \frac{d \ln \tau}{d \sigma} = N \beta \sigma^{-2}
  \label{cost1}
\end{equation}
%\begin{figure*}
%\centering
%\includegraphics[scale=0.4]{figures/fig-chi.eps} 
%\caption{Probability to transition as a function of local neighbourhood for different values of the sociality weighting $\chi$}
%\label{fig1}
%\end{figure*}
We need to find $\tau$.
%\section{Mean first passage time away from incorrect state}
Consider the process as a one-step Markov chain on the integer line from $0$ to $N$. A transition from state $j$ to $j+1$ is written $t^+(j)$ and an opposite transition is denoted $t^-(j)$. An exact expression for the mean first passage time from zero individuals correct to $N_{ESC}$ is \cite{gardiner}

@@

%\section{Transition rates}
The dynamics of an update step in stochastic decision process are as follows; first an individual is selected at random from the population, this individual has a personal estimate of the environmental cue $g_i$ and observes the opinions of all its neighbors. The decision rule of Eqn.~\ref{opt_rule} is then employed and the individual will adopt, or remain in, the state most likely to be correct.
For any value of social information $x_L$ there is a value of personal information above which the individual will switch to $S=1$. We denote this value as $g_c$ and from Eqn.~\ref{opt_rule} 
\begin{equation}
g_c  = \sigma\ln\chi  \left(\frac{1}{2}-x_L\right) 
\end{equation}
If $g_i>g_c$ then the individual will switch to $S=1$ otherwise $S=-1$. 
Dealing first with the transition rate toward the correct state, we may write this as
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^+(x) = P(S_i=-1)\sum_{x_{L}}P\left(g_i>g_c \right)P\left(x_{L}|x\right)
\label{up}
\end{equation}
where $i$ denotes the selected individual and the summation is taken over all possible values of $x_{Li}$.
where $i$ denotes the selected individual and the summation is taken over all possible values of $x_{L}$.

From Eqn.~\ref{lapdist}
\begin{align}
P\left(g_i>g_c \right) =
\frac{1}{2}\exp{\left[1/\sigma - \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }.%&\mbox{if } \\
%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%\end{cases}
\end{align}
(This expression is true if $x_L \le 1/2 - 1/(\sigma\ln\chi)$. As we are interested in these probabilities only when $x$ is small we restrict our attention to this scenario.) Given this equation we may write Eqn.~\ref{up} as
\begin{equation}
t^+(x) = (1-x)  e^{1/\sigma } \mathcal{F}(x)
\end{equation}
where 
\begin{equation}
 \mathcal{F}(x) = \frac{1}{2}\sum_{x_{L}}\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x_L\right)\right]}P\left(x_{L}|x\right)
\end{equation}
and similarly 
\begin{equation}
%t^+(x) = P(S_i=-1)\sum_{j=0}^{N_s}P\left(g_i>g_c(j/N_s)\right)P\left(x_L=j/N_s|x\right)
t^-(x) =x \left(1-  e^{1/\sigma } \mathcal{F}(x)\right)
\end{equation}
Taking the derivative of Eqn.~\ref{laplace_cost} and employing a Maclaurin series expansion leaves
\begin{equation}
  \frac{d \ln \tau}{d \sigma} \approx N \sigma^{-2} \int_{x_1}^{x_2} \left(1 + \sum_{n=1}^{\infty} \left(e^{1/\sigma } \mathcal{F}(y)\right)^n  \right) dy
\label{laplace_cost2}
\end{equation}
The extrema of the potential, $x_1,x_2$, are at the points where $t^+(x)=t^-(x)$, i.e. where $x=e^{1/\sigma } \mathcal{F}(x)$. Since $e^{1/\sigma } \mathcal{F}(x)<x$ $\forall x \in [x_1,x_2]$ the summation terms in the integrand are all strictly less than $x$, hence
\begin{equation}
  \frac{d \ln \tau}{d \sigma} \approx  N \sigma^{-2} \left({x_2}-{x_1}\right) + \mathcal{O}\left( \left(x_2-x_1\right)^2\right)
\label{laplace_cost2}
\end{equation}
Combining Eqns.~\ref{cost1} and~\ref{laplace_cost2} gives the location of minimal cost as
\begin{equation}
  \beta =  \left({x_2}-{x_1}\right) 
\label{laplace_cost_final}
\end{equation}
where in deriving this expression we have neglected terms of $\mathcal{O}(1/N)$ and $\mathcal{O}\left( \left(x_2-x_1\right)^2\right)$
Optimal strategy is therefore achieved by setting $\sigma$ so that the difference between the two extrema is equal to $\beta$.
%\begin{align}
%P\left(g_i>g_c \right) =
%1 - \frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }%&\mbox{if } \\
%%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%%\end{cases}
%\end{align}
%otherwise.
